# Chapter3

2. Lets read the alc data.
```{r}
alc <- read.csv("C:/Users/juhoi/OneDrive/Tiedostot/Coodaushit/IODS-project/create_alc.R")
names(alc)
```

3. I choose sex, final exam scores, famsize and studytime as my variables. My hypotheses are as follows:
There's less females than males who consume a lot of alcohol. 
I also assume that a person who doesn't consume a lot of alcohol does get better grades as someone who does drink a lot.

```{r}

library(tidyr)
library(dplyr)
library(ggplot2)
```

4. Now we test the hypothesis.
```{r}
alc %>% group_by(sex, high_use) %>%summarise(count=n(),mean_grade=mean(G3))
```

First here is the crosstabulation for alcohol usage, sex and the average of the final exam grades. Results tell us two things:
  i.) Higher percentage of men than women use alcohol a lot.
  ii.) For women, those who drink more have slightly higher grades. For men, binge-drinkers have worse grades.


Then we create a boxplot of high alcohol usages effects to the final grades.
```{r}
g1 <- ggplot(alc, aes(x= high_use, y=G3), col=sex)
g1 + geom_boxplot()+ylab("grade")
```
It is clear that on average, alcohol-lovers have worse grades than moderate drinkers.

Then here we have alcohol usage vs. study time as a barplot. Low drinkers in black and high drinkers in green.

```{r}
counts <- table(alc$high_use, alc$studytime)
barplot(counts, main= "Alcohol usage vs. studytime", xlab="studytime", beside=TRUE, col = c("black", "green"))
```
Key ratio here is the share of drunkards in each category of study time. The shares are about equally small in 4 and 3 but get bigger in 2 and in 1 it is almost as usual to drink a lot than not to drin a lot.

Then we compare alcohol usage to family size.
```{r}
counts2 <- table(alc$high_use, alc$famsize)
barplot(counts2, main= "Alcohol usage vs. famsize", xlab="famsize", beside=TRUE, col = c("blue", "pink"))
```
The result is that slightly higher share of students from small families, than from big families, drink a lot.

5. Logistic regression is here.
```{r}
m <- glm(high_use ~ famsize + sex + G3 + studytime, data = alc, family = "binomial")
summary(m)
```
Coefficients as odds ratios:
```{r}
OR <- coef(m) %>% exp
OR
```
COnfidence intervals:
```{r}
CI <- confint(m) %>% exp
CI
```

6. 
```{r}
library(gmodels)
m <- glm(high_use ~ sex + studytime, data = alc, family = "binomial")
probabilities <- predict(m, type = "response")
alc <- mutate(alc, probability = probabilities)
alc <- mutate(alc, prediction = probability > 0.3)

select(alc, studytime, sex, high_use, probability, prediction) %>% tail(10)
table(high_use = alc$high_use, prediction = alc$prediction)
CrossTable(alc$high_use, alc$prediction)
counts3 <- table(alc$high_use, alc$prediction)
barplot(counts3, main = "Usage of alcohol", xlab = "high_use", col = c("pink", "purple"), beside = TRUE)

g <- ggplot(alc, aes(x = probability, y = high_use, col = prediction))

g + geom_point()

table(high_use = alc$high_use, prediction = alc$prediction) %>% prop.table %>% addmargins


loss_func <- function(class, prob) {
  n_wrong <- abs(class - prob) > 0.5
  mean(n_wrong)
}

loss_func(class = alc$high_use, prob = 0)
```


According to the results, the model provided provides a sense of reality, even if it it statistically not very reliable. By simply guessing what the results would be, the ideas that one comes up with could be similar but more accurate. 
What the model probably works for though, is that is offers some slight validation for any guesses made. Not much can be concluded from the model alone however, since the ratio of wrong predictions is quite large


