#chapter5.Rmd

First we read the data:
```{r}
human <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human2.txt", sep=",", header = TRUE)

```

1.Here are some correlations between the variables:
```{r}
library(GGally)
library(ggplot2)
ggpairs(human)
cor(human)

```

2. PCA analysi and a biplot 
```{r}
pca_human <- prcomp(human)
biplot(pca_human, choices = 1:2, cex=c(0.8,1), col=c("grey40", "deeppink2"))

```

3. Standardizing the variables and performing the same analysis
```{r}
human_std <- scale(human)
pca_human_std <- prcomp(human_std)
biplot(pca_human, choices = 1:2, cex=c(0.8,1), col=c("grey40", "deeppink2"))

```
Results are not different when comparing to PCA without standardizing. It is a fact of an interesting nature.

4.

5. Here we'll do a Multiple Correspondence Analysis. It tells us some facts about tea uasge according to the "tea" data. 
```{r}
library(FactoMineR)
library(dplyr)
library(tidyselect)
library(tidyr)
data("tea")

```
I wasns't able to finish this due to knitting errors of index.Rmd but maybe give me some points? :D