# Chapter3

2. Lets read the alc data.
```{r}
alc <- read.csv("C:/Users/juhoi/OneDrive/Tiedostot/Coodaushit/IODS-project/create_alc.R")
names(alc)
```

3. I choose sex, final exam scores, famsize and studytime as my variables. My hypotheses are as follows:
There's less females than males who consume a lot of alcohol. 
I also assume that a person who doesn't consume a lot of alcohol does get better grades as someone who does drink a lot.

```{r}

library(tidyr)
library(dplyr)
library(ggplot2)
```

4. Now we test the hypothesis.
```{r}
alc %>% group_by(sex, high_use) %>%summarise(count=n(),mean_grade=mean(G3))
```

First here is the crosstabulation for alcohol usage, sex and the average of the final exam grades. Results tell us two things:
  i.) Higher percentage of men than women use alcohol a lot.
  ii.) For women, those who drink more have slightly higher grades. For men, binge-drinkers have worse grades.


Then we create a boxplot of high alcohol usages effects to the final grades.
```{r}
g1 <- ggplot(alc, aes(x= high_use, y=G3), col=sex)
g1 + geom_boxplot()+ylab("grade")
```
It is clear that on average, alcohol-lovers have worse grades than moderate drinkers.

Then here we have alcohol usage vs. study time as a barplot. Low drinkers in black and high drinkers in green.

```{r}
counts <- table(alc$high_use, alc$studytime)
barplot(counts, main= "Alcohol usage vs. studytime", xlab="studytime", beside=TRUE, col = c("black", "green"))
```
Key ratio here is the share of drunkards in each category of study time. The shares are about equally small in 4 and 3 but get bigger in 2 and in 1 it is almost as usual to drink a lot than not to drin a lot.

Then we compare alcohol usage to family size.
```{r}
counts2 <- table(alc$high_use, alc$famsize)
barplot(counts2, main= "Alcohol usage vs. famsize", xlab="famsize", beside=TRUE, col = c("blue", "pink"))
```
The result is that slightly higher share of students from small families, than from big families, drink a lot.

5. Logistic regression is here.
```{r}
m <- glm(high_use ~ famsize + sex + G3 + studytime, data = alc, family = "binomial")
summary(m)
```
Coefficients as odds ratios:
```{r}
OR <- coef(m) %>% exp
OR
```
COnfidence intervals:
```{r}
CI <- confint(m) %>% exp
CI
```

6. Sex ja study time seem to have a statistical relationship to alcohol use. That's why we modify the model to include only sex and study time.
```{r}
m <- glm(high_use ~ sex + studytime, data = alc, family = "binomial")
```
Then we'll crosstabulate the 


