# Chapter2
The wrangling was very exciting but I hope analysis part consists of something more concrete.

First, we read the analysis_dataset data and get these additional functions that I'm going to use later.

```{r}
analysis_dataset <- read.table("C:/Users/juhoi/OneDrive/Tiedostot/Coodaushit/IODS-project/data/learning2014.txt", header = TRUE, sep = " ")
library(GGally)
library(ggplot2)
```


1.
The analysis_dataset is a set of 166 observations from 7 variables. Variables are the age and gender of participants of the study as well as thei points in an exam and their attitude towards statistics. Also in the variables, there are three sets: Stra(tegic) meaning the questions related to studying strategies, Deep meaning the questions related to long-term studying techniques and Surf(ace) meaning the questions related to questions related to short-term sudying techniques.

2. Now we visualize the analysis_dataset data.


```{r}
p <- ggpairs(analysis_dataset, mapping = aes(col = gender, alpha = 0.3), lower = list(combo = wrap("facethist", bins = 20)))
p
```


Here we have a rather complex plot. it gives us information on the relations between the variables. For example in low left corner (box (1,1); as in coordination) we can see a distribution of exam points within each gender.In that box we can clearly see that there was more variance in the points of women than in the points of men. Other interesting notions could be the difference between attitudes towards statistics between men and women in the box (3,5). Scatterplots such as in box (2,5), are a little harder to interpret but there might be a correlation between higher age and more negative attitude towards statistics in the box(2,5).

Then we'll print a summary of the analysis_dataset
summary(analysis_dataset)

This tells us some interesting data of each variable respectively. For example that the mean of the ages is 25.51 years. And that almost twice as much women than men took part to the survey.Basically all the pieces of information that the summary gives are related to the distribution of each variable.


3. Then we'll do a regression model using points as the target variable. I choose age, attitude and deep as explanatory variables. Also below is the summary of the fitted model.
```{r}
my_model <- lm(formula = points ~ attitude + gender + surf, analysis_dataset)
my_model

summary(my_model)
```


This kind of linear regression is usually used to find out what causes something. For example how big of a role does unemployment benefit play in the quantity unemployed people. The output of the model consists of coefficients which are rates with wich each of the explanatory variables affect the target variable. 
The summary function gives more information about how reliable the coefficients are. 

4. A quick look at the results tells that all the explanatory variables are statistically significant since their p-values are over 0,05 and their affect to the target variable is as follows:
Quite naturally having a good attitutde towards statistics, causes significantly better exam results.
Being a female makes it slightly more likely that you'd have worse exam results than you male peers.
Then, having higher score in surface-questions, increases likelyhood to have worse exam results than peers with lower surface-score. This is also a pretty obvious result given the nature of surface-questions.

Multiple squared of this regression is 0,197. This symbols the distance between the points and the regression line. Linear regression is alltogether basically just minimizing this distance.

5. Here are three diagnostics plots of this regression. 

```{r}
plot(my_model, which=c(1,2,5))
```

The residuals vs. fitted plot speaks for the validity of the regression. In a reliable regression the points should be scatterered randomly and the line in the center about straight from right to left. This is roughly the case here.

Normal Q-Q plot looks also very good. Some extreme values are a bit of the line though.

Residuals vs. leverage tells which observations are not only extreme outliers but also significant. In this case there are not a lot points in far right corners so also this plot suggests that the regression is reliable.
