#Chapter6
Juho Ivaska, 9.12.2018

This final peace of course work is named chapter 6  because I thought it would be logical. We use the BPRS and RATS data from the previously done wrangling.

First, some packages maybe or maybe not needed in the future:
```{r}
library(data.table)
library(dplyr)
library(stringr)
library(ggplot2)
library(GGally)
library(tidyr)
```

1.

First we shall work with RATS data and do some plots. I had some troubles with knitting this file so I put all the wrangling done before here to make sure knitting goes well.
```{r}
BPRS <- read.table("https://raw.githubusercontent.com/KimmoVehkalahti/MABS/master/Examples/data/BPRS.txt", header = TRUE, sep = " ")
RATS <- read.table("https://raw.githubusercontent.com/KimmoVehkalahti/MABS/master/Examples/data/rats.txt", header = T, sep = "\t")

BPRS$treatment <- factor(BPRS$treatment)
BPRS$subject <- factor(BPRS$subject)

RATS$ID <- factor(RATS$ID)
RATS$Group <- factor(RATS$Group)

BPRSL <- BPRS %>%
  gather(week, BPRS, week0:week8)
RATSL <- RATS %>%
  gather(Time, Weight, WD1:WD64)
```


Here's a plot which gives the change of rats' weights through the study period of 9 weeks. In the study, three groups of rats were fed differently and scientists followed their weight.
```{r}
ggplot(RATSL, aes(x = Time, y = Weight, group = ID, linetype = Group)) +
  geom_line()


```

We can see that all the weights followed quite steadily the same path with some outliers. They all also seem to have a loss in the end. Three different line types symbol the three different group of rats.

Now here is a boxplot of the means of the weights of the different groups of mice:
```{r}
RATSL64 <- RATSL %>%
  filter(Time > 1) %>%
  group_by(Group, ID) %>%
  summarise(mean = mean(Weight)) %>%
  ungroup()

ggplot(RATSL64, aes(x = Group, y = mean)) +
  geom_boxplot() +
  stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "mean(weight), days 1-64")
```

We can see that the mean is the lowest with the group one.

Then we want to show the same boxplots wihtout some outliers. This should give more accurate results.
```{r}
RATSL641 <- RATSL %>%
  filter(Time > 1, Weight < 580) %>%
  group_by(Group, ID) %>%
  summarise(mean = mean(Weight)) %>%
  ungroup()

ggplot(RATSL641, aes(x = Group, y = mean)) +
  geom_boxplot() +
  stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "mean(weight), days 1-64")

```

Then we'll move to analyze BPRS data. 
First we'll make on boxplot to see how BPRS scores developed through the study period with each treatment
```{r}
ggplot(BPRSL, aes(x = week, y = BPRS, col=treatment)) +
  geom_boxplot()
```

Next we can see a linear model made to notice differencies between the two treatments:
```{r}
BPRS_reg <- lm(BPRSL$BPRS ~ BPRSL$treatment)
summary(BPRS_reg)
```

And here's is a Random intercept model which allows more formal analysis of the data:
```{r}
library(lme4)

BPRSL_ref <- lmer(BPRS ~ treatment + (1 | subject), data = BPRSL, REML = FALSE)
summary(BPRSL_ref)
```

Now we have taken a nice look to the data and made some very decent notions. Nice work!